Biology background

### 2014-12-10 Rmarkdown/biology_background.Rmd
Edits to biology_background

Sir Adrian Bird, an esteemed British geneticist and someone who is in favour of the broader definition, attempts to unite these definitions in the following \citep{Bird:2007em}:

@@

DNA methylation is a chemical modification of DNA that can impart information on top of the DNA sequence. It is heritable during mitotic cell division, which means that it is faithfully copied across to the daughter cell during cell division[^dnam_copying]. It therefore fits into Arthur Riggs' afore-mentioned definition of an epigenetic modification \citep[p1]{russo1996epigenetic}. DNA methylation is found in bacteria, fungi, plants and animals.

@@

1. _Enzyme digestion_
2. _Affinity enrichment_
3. _Sodium bisulfite conversion_

@@

Almost all assays of DNA methylation measure a "population average" from a pool of hundreds or thousands of cells. For a diploid organism, this is an average over several different levels: the two DNA strands, the two homologous chromosomes within a diploid cell and the hundreds or thousand of cells used in the assay. Hundreds or thousands of cells are required in order to have sufficient material as input for the assay. Assays that require only a single cell as input do exist \citep[e.g.,]{Smallwood:2014kn, Guo:2013ih} but are still in development and not yet in widespread use.

@@

Some examples of these targeted bisulfite-sequencing are:
- Reduced representation bisulfite-sequencing (_RRBS_, \citet{Meissner:2005eb}): RRBS uses restriction enzymes to first select regions of the genome with a high CpG density (based on the Msp-I cleavage motif), which are subsequently treated with sodium-bisulfite and sequenced.
- Extended reduced representation bisulfite-sequencing, also known as enhanced reduced representation bisulfite-sequencing (eRRBS, \citet{Akalin:2012he}). A modified version of the RRBS protocol.
- NimbleGen's SeqCap Epi Enrichment System and Agilent's SureSelectxt Human Methyl-Seq: Both of these commercial products use DNA hybridisation to enrich the sequencing library for pre-defined regions of interest. This enriched library is then bisulfite-converted and sequenced.

>>>>

Sir Adrian Bird, an esteemed British geneticist, and someone who is in favour of the broader definition, attempts to unite these definitions in the following \citep{Bird:2007em}:

@@

DNA methylation is a chemical modification of DNA that can impart information on top of the DNA sequence. It is heritable during mitotic cell division, which means that it is faithfully copied across to the daughter cell during cell division[^dnam_copying]. It therefore fits into Arthur Riggs' aforementioned definition of an epigenetic modification \citep[p1]{russo1996epigenetic}. DNA methylation is found in bacteria, fungi, plants and animals.

@@

1. Enzyme digestion
2. Affinity enrichment
3. Sodium bisulfite conversion

@@

Almost all assays of DNA methylation measure a population average from a pool of hundreds or thousands of cells. For a diploid organism, this is an average over several different levels: the two DNA strands, the two homologous chromosomes within a diploid cell and the hundreds or thousand of cells used in the assay. Hundreds or thousands of cells are required in order to have sufficient material as input for the assay. Assays that require only a single cell as input do exist \citep[e.g.,]{Smallwood:2014kn, Guo:2013ih} but are still in development and not yet in widespread use.

@@

Some examples of these targeted bisulfite-sequencing are:

- Reduced representation bisulfite-sequencing (_RRBS_, \citet{Meissner:2005eb}): RRBS uses restriction enzymes to first select regions of the genome with a high CpG density (based on the Msp-I cleavage motif), which are subsequently treated with sodium-bisulfite and sequenced.
- Extended reduced representation bisulfite-sequencing, also known as enhanced reduced representation bisulfite-sequencing (eRRBS, \citet{Akalin:2012he}). A modified version of the RRBS protocol.
- NimbleGen's SeqCap Epi Enrichment System and Agilent's SureSelectxt Human Methyl-Seq: Both of these commercial products use DNA hybridisation to enrich the sequencing library for pre-defined regions of interest. This enriched library is then bisulfite-converted and sequenced.
